---
format:
  gfm:
    default-image-extension: ""
---

<!-- README.md is generated from README.qmd. Please edit that file -->

<!-- Rendering runs the examples, which build the 2017 data (about 230 MB of downloads) unless they are already cached. -->

```{r}
#| include: false
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
options(pillar.width = 80)
# Keep the start-up messages of duckplyr and DuckDB out of the README.
suppressMessages(dplyr::collect(duckplyr::duckdb_tibble(x = 1)))
```

# freightdataus

<!-- badges: start -->
[![R-CMD-check](https://github.com/UchidaMizuki/freightdataus/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/UchidaMizuki/freightdataus/actions/workflows/R-CMD-check.yaml)
[![Codecov test coverage](https://codecov.io/gh/UchidaMizuki/freightdataus/graph/badge.svg)](https://app.codecov.io/gh/UchidaMizuki/freightdataus)
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
<!-- badges: end -->

freightdataus provides ready-to-use US freight flow data from the Commodity Flow Survey (CFS) Public Use File for 2012, 2017, and 2022.
It is meant as ground-truth data for validating methods that estimate origin-destination flows.

The data are downloaded from the US Census Bureau and built the first time you request them, using [tarchives](https://github.com/UchidaMizuki/tarchives).
Later requests read the cached results.

## Installation

You can install the development version of freightdataus from [GitHub](https://github.com/) with:

``` r
# install.packages("pak")
pak::pak("UchidaMizuki/freightdataus")
```

## Usage

```{r}
#| label: setup
library(freightdataus)
```

Weighted flows between states or CFS Areas, by commodity and optionally by mode:

```{r}
#| label: flow
cfs_flow_get(2017, geography = "cfs_area", by_mode = TRUE)
```

Zones with their centers of population, their boundaries (an sf data frame), and the distances between zones:

```{r}
#| label: zone
cfs_zone_get(2017, geography = "cfs_area")
cfs_pair_get(2017, geography = "cfs_area")
```

```{r}
#| label: zone-boundary
#| eval: false
cfs_zone_boundary_get(2017, geography = "cfs_area")
```

Shipment records, returned lazily as a duckplyr frame, to tabulate them in other ways:

```{r}
#| label: shipment
cfs_shipment_get(2017) |>
  dplyr::summarise(
    value_usd = sum(weighting_factor * value_usd),
    .by = mode_code
  ) |>
  dplyr::arrange(mode_code) |>
  dplyr::collect()
```

Inside a [targets](https://docs.ropensci.org/targets/) pipeline, declare targets that read the data:

``` r
# _targets.R
library(targets)

list(
  freightdataus::cfs_flow_target(flow, year = 2017, geography = "cfs_area"),
  freightdataus::cfs_pair_target(pair, year = 2017, geography = "cfs_area")
)
```

Every year returns the same columns, with the same types and units.
Weights are in metric tonnes and distances in km (`units` columns), and values are in US dollars.
Weighted totals are the sum of the weighting factor times the variable, as the CFS users guides specify.

Codes are kept as published, and some of them differ between years.
In particular, the mode codes of 2022 are not those of 2012 and 2017, and CFS Area boundaries change between survey years.
